NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-29 paper/task.tex
update 6.1

\section{Task Analysis}
\label{sec:task}
Identify the analysis goals and the corresponding visualization tasks is the first yet significant step in designing a useful tool. In this section, we examine how domain experts normally approach the task of analyzing a model (i.e., assess its strength and weakness) and identify where and how an interactive visualization system can aid in such a process.
Then, we distill the numerous discussions with domain experts into concrete tasks, which guide the design and implementation of the proposed tool.

During our long-term collaboration with NLP experts, we have conducted extensive discussion to understand the common approach employed by researcher for assessing the behavior of a model.%
The prediction accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model, however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pickup unintended pattern in the training data that can not be generalized in real word scenarios).
%
As a result, the domain experts rely on an exploration centric approach to conduct error analysis and obtain intuition.

There is little one can learn when everything is behaving as expect, therefore,
observing how models make mistakes under various scenarios is at the center of such a process.
The experts often start with simple failure examples, and then make minor perturbations (replace a word or phrase) to the input and observe the change in the prediction. This exercise helps isolate where in the input is likely contributing to the failures. For many NLP models, the attention information (see Section~\ref{sec:attention}) is essential to infer the mechanism of the model. Experts often print out the attention values or generate plots to visualize sentence alignment or compare attention values. As the experts explore more variation of similar examples, combined with their domain knowledge, he / she may develop hypotheses about the cause of failures, or ask additional questions, which lead to further experimentation.

@@

%Through many rounds of discussions with NLP experts,
%In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%
To support the visual exploratory workflow, we have summarized the following visualization tasks.
>>>>>>> More small changes
%
To support the visual exploratory workflow, we need to address the following visualization tasks.

@@

\item To begin with, understanding how perturbation (e.g., replacing words with synonymous) of the input affect the prediction is the curial part of error analysis. It not only measures the prediction sensitivity, but also help reveal the potential source of the errors in the input (\textbf{T1}).

%Understand how attention affect the prediction.
\item In addition, study the role and mechanism of attention by examine how perturbation in the input sentence affect the attention, and how changes in attention affect prediction (\textbf{T2}).
%Study the relationship between model stages (encoder, attention, classifier) and the prediction

\item So far, we focus on how changes in earlier stage of the model affect later stages or final prediction. However, what if the current prediction is wrong, how can we apply minimal update to the model to produce correct prediction. and how would such a modification affect attention (\textbf{T3}).
% Therefore, thirdly, we want to explore how model should be updated to produce the desirable change of prediction, and how the attention respond to such a change (\textbf{T3}).

>>>>

\section{Task Analysis}
\label{sec:task}
Identifying the analysis goals and the corresponding visualization tasks is the first significant step in designing a useful tool. In this section, we examine how domain experts normally approach the task of analyzing a model (i.e., assess its strength and weakness) and identify where and how an interactive visualization system can aid in such a process.
Then, we distill the numerous discussions with domain experts into concrete tasks, which guide the design and implementation of the proposed tool.

During our long-term collaboration with NLP experts, we have conducted extensive discussions to understand the common approach employed by researcher for assessing the behavior of a model.%
The predictive accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model; however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pick up an unintended pattern in the training data that cannot be generalized in real-word scenarios).
%
As a result, the domain experts rely on an exploration-centric approach to conduct error analysis and obtain intuition.

There is little one can learn when everything is behaving as expected, and therefore,
observing how models make mistakes under various scenarios is at the center of such a process.
The experts often start with simple failure examples, and then make minor perturbations (replace a word or phrase) to the input and observe the change in the prediction. This exercise helps isolate where in the input likely contribute to the failures. For many NLP models, the attention information (see Section~\ref{sec:attention}) is essential to infer the mechanism of the model. Experts often print out the attention values or generate plots to visualize sentence alignment or compare attention values. As the experts explore more variation of similar examples, combined with their domain knowledge, they may develop hypotheses about the cause of failures, or ask additional questions that lead to further experimentation.

@@

%Through many rounds of discussions with NLP experts,
%In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%
To support the visual exploratory workflow, we need to address the following visualization tasks.

@@

\item To begin with, understanding how perturbation (e.g., replacing words with synonymous) of the input affects the prediction is the crucial part of error analysis. It not only measures the prediction sensitivity, but also helps reveal the potential source of the errors in the input (\textbf{T1}).

%Understand how attention affect the prediction.
\item In addition, we study the role and mechanism of attention by examining how perturbation in the input sentence affect the attention, and how changes in attention affect prediction (\textbf{T2}).
%Study the relationship between model stages (encoder, attention, classifier) and the prediction

\item So far, we focus on how changes in earlier stage of the model affect later stages or the final prediction. However, what if the current prediction is wrong, how can we apply minimal update to the model to produce correct prediction. and how would such a modification affect attention (\textbf{T3}).
% Therefore, thirdly, we want to explore how model should be updated to produce the desirable change of prediction, and how the attention respond to such a change (\textbf{T3}).
